NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-08-26 paper/Chapters/Chapter3.tex
abstract done
grammarly checked

A series of traditional LCA of the responses to carbohydrate intake within 7 time slots of day was first examined. These initial analyses ignored the clustering of observation days within participants of the survey. \textbf{Table \ref{tab:mixmodels}} shows the latent class solutions for one to five classes (see rows under the Fixed effects model section). The BIC declines with the number of day level classes increases. However, the improvement of BIC dropped to less than 1000 from 3 classes to 4 classes solutions (658.9) and from 4 classes to 5 classses solutions (361.7). Entropy index indicates that the 4 classes model could explain about 51\% percent of the data, while \textit{p} values of Lo-Mendell-Rubun LRT suggest that the more classes we fit, the better model we will have until up to 6 classes (\textit{p} = 0.06 and is not shown in the table). From the parsimony point of view, we extended the model with random effects building on 2 classes, 3 classes and 4 classes solutions.

The results of the random effect included models are presented in \textbf{Table \ref{tab:mixmodels}} under the Random effects model section. It is obvious that the BIC improves with the addition of the random effects which account for the nested structure of the data. Entropy indicates that 2 classes in the day level and 4 classes in the individual level may be the best solution mathematically. However, after these solutions were checked in more details, the potentially most substantively interpretable model was found to be the 3$\times$3 random effect model, which is the model with 3 latent classes in the day level, and 3 latent classes in the individual level. We must emphasize that different researchers may have made decision slightly different from ours, we have provided other solutions in the \textbf{Appendix \ref{AppendixC}} for reference.

In the 3$\times$3 random effect model solution we have chosen, there were 39.5\%, 20.4\%, and 40.1\% observations classified into 3 latent groups in the day level. The overall counts and percentages for each responses within each time slot and the distributions of the solution are presented in \textbf{Table \ref{tab:daylevel}}. The trajectories illustrating the change of the probabilities of each response to carbohydrate eating during the hours of the day are shown separately by three types of days in \textbf{Figure \ref{fig:level1}}.

@@

Class 1 days \textbf{(Figure \ref{fig:level1}-A)} were given the name of "high percentage carbohydrate day" since in these days of survey, the probabilities of carbohydrate contributed higher or equal to 50\% of the energy consumed were always higher than that in the other two types of days. Specifically, high percentage carbohydrate days were characterised with probabilities of over 0.6 in time slots between 6 am to 9 am, 9 am to 12 noon, and also 2 pm to 5 pm, during which the time slots may be interpreted as breakfast, morning snack, and afternoon snack time periods for many participants. Moreover, even during late night time period, such as 8 pm to 10 pm, and 10 pm to 6 am time slots, the probabilities of having higher carbohydrate contained food were still as high as 0.412, and 0.246, respectively.

@@

Class 2 days \textbf{(Figure \ref{fig:level1}-B)} were named as "low percentage carbohydrate day" because first of all, in these days the probability of participants skipping breakfast was 0.45. And after 9 am, within these days, the probability of having low carbohydrate contained food (carbohydrate contributed $<$ 50\% of total energy intake), was always higher than having high carbohydrate contained food (carbohydrate contributed $\geqslant$ 50\% of total energy intake). In class 2 days, participants also turned to have morning snacks (with only 0.079 possibility of \textbf{not} eating any food and similar probabilities of having either high or low carbohydrate contained food). This phenomenon may also be interpreted as having a long and late breakfast (brunch) in these mornings. The probability of \textbf{not} eating any food was the lowest for low carbohydrate days during the midnight time slot (10 pm to 6 am), with probability of 0.590 compared with 0.680 and 0.751 in class 1 and class 3 days, respectively.

@@

As expected, consumption of total energy (7539.98 kJ), total carbohydrate (222.79 g), total sugar (98.63 g), starch (124.07 g), and non-milk extrinsic sugar (59.45 g) were the highest among high percentage carbohydrate days (all \textit{p} < 0.001). On the other hand, the consumption of protein (73.29 g), total fat (67.24 g), and alcohol (13.80 g) were the highest in the regular meals days. Moreover, in high percentage carbohydrate days, participants turned to consume the highest amount of fruit (107.40 g). There was no evidence of any difference for the consumption of yellow, red, or green vegetables across the three types of days (\textit{p} = 0.681).\vspace{-0.6cm}

@@

\subsection{Individual level LCA solution}\vspace{-0.3cm}

In the random effect models we utilized the non-parametric approach, in which we added a level 2 (individual level) latent classes based on the random means from the level 1 (day level) latent class solution. The results of the individual level LCA solution for 2 and 3 classes are presented in \textbf{Figure \ref{fig:CB2level2}}, and \textbf{\ref{fig:level2}}.

@@

With two individual level latent classes \textbf{(Figure \ref{fig:CB2level2})}, one individual class is comprised of individuals with a relatively slightly higher proportion of having "low carbohydrate day" (22.1\%) compared to the other (17.4\%). This class represents nearly 65\% of the individuals. However, we believe these individual classes are not very distinguishable to each other.

With three individual level latent classes \textbf{(Figure \ref{fig:level2})}, a low-carbohydrate eaters class, a moderate-carbohydrate eaters class, and a high-carbohydrate eaters class emerges. 43.1\% participants were identified as high-carbohydrate eaters, in these individuals, about 50\% of the days (2 out of 4 days) of their dietary diary could be classified as having high percentage carbohydrate days. Nearly 1 out of 4 days of their dietary diary were either "regular meals day" or "low carbohydrate day". 28.1\% participants fell into the low carbohydrate eaters class in the left hand side of \textbf{Figure \ref{fig:level2}}, their recordings of food intake showed that in more than 60\% of their days, they were having "regular meals" which was characterised as with highest amount of fat and alcohol consumptions as already described in \textbf{Table \ref{tab:day-level-features}}. Moderate carbohydrate eaters have comparable proportions (42.0\% vs. 40.0\%) of having high carbohydrate days and regular meals day, 18.0\% of their dietary diary were found to be low carbohydrate days.

@@

After recognising that there were three potential latent groups of carbohydrate eaters in the UK adults, whose food consumption pattern were also probably switching from one to another during the survey, their average carbohydrate contribution to total energy intake (as well as the subtypes of carbohydrate actually consumed) within the 7 pre-defined time slots of the day were still of interest. Survey-design-weighted mean energy intake within each time slot of the day and their composition of contribution are illustrated in \textbf{Figure \ref{fig:energysourcesCB}}, weighted mean nutrients intakes are listed in \textbf{Table \ref{tab:tab-nutri-indi}}.

@@

Among the three types of carbohydrate eaters, the mean of total energy intake over the 4 days of dietary survey was the highest (7985.8 kJ, 95\%CI: 7283.3, 8146.3) in the low carbohydrate eaters group, and the lowest (7341.8 kJ, 95\%CI: 7172.5, 7511.2) in the moderate eaters group \textbf{(Table \ref{tab:tab-nutri-indi})}. Sources of energy for each type of carbohydrate eaters by the 7 time slots were also different. Low carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-A)} never had carbohydrate contributed more than 50\% of their total energy throughout the day. Energy from fat were the highest for low carbohydrate eaters most of the time during the day (except for time between 10 pm to 6 am next morning). Most impressively, energy contributed by alcohol was always the highest in low carbohydrate eaters, percentages for energy from alcohol for the 7 times slots were 1.4\% (6-9 am), 1.8\% (9-12 noon), 3.1\% (12-2 pm), 7.2\% (2-5 pm), 8.4\% (5-8 pm), 20.0\% (8-10 pm), and 31.2\% (10-6 am), respectively. Contributions from different energy sources were quite similar for moderate and high carbohydrate eaters, but their absolute amount of energy consumption at each time slot were largely different. Moderate carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-B)} were characterised as consuming the lowest energy (296.9 kJ) before 9 am, but having higher energy consumption (967.9 kJ) between 9 am and 12 noon time compared with low and high carbohydrate eaters. Moderate carbohydrate eaters may turn to have later breakfast, later lunch, and probably later dinner as well. They had the highest total energy consumption (599.7 kJ) at night (10 pm - 6 am) across three types of eaters. High carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-C)} consumed the highest total energy (929.0 kJ) during 6 am to 9 am in the morning and the lowest total energy between 10 pm to 6 am (205.5 kJ). Specifically, carbohydrate contributions to total energy intake were 65.3\% (6-9 am), 56.2\% (9-12 noon), 46.2\% (12-2 pm), 50.3\% (2-5 pm), 42.5\% (5-8 pm), 44.1\% (8-10 pm), and 49.9\% (10-6am). We also noticed that high carbohydrate eaters consumed their energy mainly from three time slots: 6-9 am, 12-2 pm, and 5-8 pm.

As expected, in total, the mean of total carbohydrate intake was 203.8 g, 218.3 g, and 233.4 g for low, moderate, and high carbohydrate eaters, respectively \textbf{(Table \ref{tab:tab-nutri-indi})}. Total energy contributed by carbohydrate was close to 50\% in the high carbohydrate eaters, but was only 40.6\% in the low carbohydrate eaters. In terms of the subtypes (components) of the carbohydrate consumed at each time slot, high carbohydrate eaters consumed as much as more than 2 times (compared to low carbohydrate eaters) and nearly 4 times (against moderate carbohydrate eaterss) the amount of sugar (37.9g 95\% CI: 36.8, 39.2) and non-milk extrinsic sugar (i.e. free sugar) (11.1g 95\%CI: 10.7, 11.6) between 6-9 am. Moderate carbohydrate eaters had their carbohydrate intake more spread out. They consumed more sugar and starch during 9-12 noon, 2-5 pm, 8-10 pm, and 10-6 am. Low carbohydrate eaters turned to have similar temporal pattern of consuming carbohydrates but the absolute amount of fibre, sugar, free sugar, and starch were usually lower than that in the high carbohydrate eaters except for time slots of 2-5 pm, and 10-6 am. Strong evidence (\textit{p} < 0.001) suggested that the mean of total fibre consumption for low,

@@

moderate, and high carbohydrate eaters were different: 13.7g (13.4, 14.0), 12.5g (12.1, 12.9), and 14.7g (14.4, 14.9) with all 95\% CI being exclusive to each other. It is also noteworthy that low carbohydrate eaters consumed the highest average amount of protein (79.9 g, 17.2\% of total energy), fat (74.7g, 35.4\% of total energy), and alcohol (20.8 g, 6.8\% of total energy) as we have described for \textbf{Figure \ref{fig:energysourcesCB}}.

The social-demographic characteristics of the UK adults according to their individual level latent class membership are shown in \textbf{Table \ref{tab:Level2tab1}}. Moderate carbohydrate eaters were relatively younger (\textit{p} < 0.001). Gender distribution across the three types of carbohydrate eaters was fairly even (\textit{p} = 0.119). The distribution of the carbohydrate eater types turned out to be changing with the year of survey. Low carbohydrate eaters represented 32.5\% of the population in the first year of survey, but later dropped to lower than 30\% (lowest in the third year, 22.6\%) until the most recent year. Proportion of high carbohydrate eaters increased from 41.2\% to the highest (50.6\%) in the second year, but then started to decline to 38.4\% in the 8th year of survey (\textit{p} = 0.015). There was no evidence of difference in employment status across three types of carbohydrate eaters. However, strong evidence suggested that high carbohydrate eaters had the highest proportion (61.3\%) of living with partner (\textit{p} < 0.001); moderate carbohydrate eaters had the lowest average income (27180.8 \textsterling/year), highest proportion of non-white population (20.5\%), and lower education level (23.3\% with degree of higher education) compared with either low or high carbohydrate eaters.

@@

\caption{\label{tab:Level2tab1}Weighted means, percentages, and
95\% CIs of the social-demographic characteristics by carbohydrate eating latent class memberships in the UK
adults. \\ (NDNS RP 2008/09-15/16, sample size = 6155)} \centering
\fontsize{9}{11}\selectfont

\begin{tabular}[t]{lcccc}
	\hiderowcolors
	\toprule
	\textbf{Variables} & \textbf{\Centerstack{Low carbo-\\hydrate eaters\\(n = 1730)}} & \textbf{\Centerstack{Moderate carbo-\\hydrate eaters\\(n = 1772)}} & \textbf{\Centerstack{High carbo-\\hydrate eaters\\(n = 2653)}} & \textbf{\textit{P} value} \textsuperscript{*}\\
	\midrule
	\showrowcolors
	Total (\%) & 28.4 (26.9, 29.9) & 28.7 (27.1, 30.3) & 43.0 (41.3, 44.7) & \\
%	Country (\%) &  &  &  & 0.007\\
%	\hspace{1em}England & 84.5 (81.7, 86.9) & 82.0 (79.3, 84.5) & 84.7 (82.3, 86.8) & \\
%	\hspace{1em}Northern Ireland & 2.1 (1.6, 2.8) & 4.2 (3.2, 5.6) & 2.2 (1.7, 3.0) & \\
%	\hspace{1em}Scotland & 9.1 (7.0, 11.8) & 8.6 (6.7, 11.1) & 8.0 (6.3, 10.2) & \\
%	\hspace{1em}Wales & 4.3 (3.3, 5.6) & 5.1 (4.0, 6.4) & 5.1 (4.0, 6.4) & \\
	Age (years) & 51.0 (49.9, 52.1) & 40.3 (39.1, 41.6) & 51.7 (50.7, 52.7) & < 0.001\\
	Sex (\%) &  &  &  & 0.119\\
	\hspace{1em}Men & 50.0 (46.9, 53.1) & 50.2 (47.0, 53.5) & 46.6 (44.0, 49.1) & \\
	\hspace{1em}Women & 50.0 (46.9, 53.1) & 49.8 (46.5, 53.0) & 53.4 (50.9, 56.0) & \\
	Survey years (\% in rows) &  &  &  & 0.015\\
	\hspace{1em}1 & 32.5 (28.4, 36.9) & 26.3 (21.9, 31.2) & 41.2 (36.6, 46.0) & \\
	\hspace{1em}2 & 26.8 (22.6, 31.3) & 22.6 (18.6, 27.3) & 50.6 (45.8, 55.4) & \\
	\hspace{1em}3 & 22.6 (18.8, 26.9) & 33.7 (28.6, 39.2) & 43.6 (38.7, 48.7) & \\
	\hspace{1em}4 & 27.9 (24.1, 32.2) & 27.6 (23.8, 31.8) & 44.4 (40.2, 48.7) & \\
	\hspace{1em}5 & 27.9 (24.2, 32.0) & 28.7 (24.4, 33.5) & 43.3 (38.2, 48.6) & \\
	\hspace{1em}6 & 28.0 (24.0, 32.4) & 31.5 (26.9, 36.6) & 40.5 (35.8, 45.3) & \\
	\hspace{1em}7 & 29.1 (25.2, 33.4) & 29.0 (24.5, 34.0) & 41.8 (37.1, 46.7) & \\
	\hspace{1em}8 & 31.1 (27.3, 35.3) & 30.5 (25.9, 35.5) & 38.4 (34.1, 42.8) & \\
	Paid employment\textsuperscript{\dag} (\%) &  &  &  & 0.907\\
	\hspace{1em}Yes & 40.3 (37.0, 43.6) & 40.8 (37.1, 44.5) & 39.8 (37.1, 42.6) & \\
	\hspace{1em}No & 59.7 (56.4, 63.0) & 59.2 (55.5, 62.9) & 60.2 (57.4, 62.9) & \\
	Live with partner\textsuperscript{\ddag} (\%) &  &  &  & < 0.001\\
	\hspace{1em}Yes & 56.9 (53.6, 60.1) & 38.4 (35.2, 41.8) & 61.3 (58.7, 63.7) & \\
	\hspace{1em}No & 43.1 (39.9, 46.4) & 61.6 (58.2, 64.8)) & 38.7 (36.3, 41,3) & \\
	Household income, \textsterling/year & \Centerstack{36558.5\\(34800.2, 38316.8)} & \Centerstack{27180.8\\(25597.9, 28763.7)} & \Centerstack{32171.6\\(31024.9, 33318.2)} & < 0.001\\
	Ethnicity (\%) &  &  &  & \\
	\hspace{1em}White & 94.2 (92.4, 95.6) & 79.5 (76.4, 82.3) & 91.9 (90.1, 93.4) & < 0.001\\
	\hspace{1em}Non-White & 5.8 (4.4, 7.6) & 20.5 (17.7, 23.6) & 8.1 (6.6, 9.9) & \\
	Education (\%) &  &  &  & \\
	\hspace{1em}Degree or higher & 29.0 (26.1, 32.1) & 23.3 (20.5, 26.3) & 26.2 (24.1, 28.5) & 0.019\\
	\hspace{1em}Lower than degree & 71.0 (67.9, 73.9) & 76.7 (73.7, 79.5) & 73.8 (71.5, 75.9) & \\
	\bottomrule
	\multicolumn{5}{l}{{\scriptsize \textit{Note: }}}\\
	\multicolumn{5}{l}{{\scriptsize \textbf{Abbreviations}: CI, confidence intervals; NDNS RP, national diet and nutrition survey rolling programme.}}\\
	\multicolumn{5}{l}{{\scriptsize Variables were weighted by individual weights.}}\\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{*} For continuous variables, the \textit{F} test was used to determine differences between latent classes with Bonferroni}}\\ 
	\multicolumn{5}{l}{{\scriptsize correction to account for multiple testing across $>$ 2 classes. For categorical variables, differences between}}\\
	\multicolumn{5}{l}{{\scriptsize  latent classes were assessed using the adjusted Pearson $\chi^2$ test for survey data.}} \\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{\dag} Paid employment was defined as being in paid employment during the last 4 weeks prior to the survey.} }\\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{\ddag} Live with partner was defined as either living with a married husband/wife or a legally recognised civil}} \\
	\multicolumn{5}{l}{{\scriptsize partnership.}}\\
\end{tabular}

@@

\subsection{Association between individual level latent classes and hypertension, and obesity.}\vspace{-0.3cm}

>>>>

A series of traditional LCA of the responses to carbohydrate intake within 7-time slots of the day was first examined. These initial analyses ignored the clustering of observation days within participants of the survey. \textbf{Table \ref{tab:mixmodels}} shows the latent class solutions for one to five classes (see rows under the Fixed effects model section). The BIC declines with the number of day level classes increases. However, the improvement of BIC dropped to less than 1000 from 3 classes to 4 classes solutions (658.9) and from 4 classes to 5 classes solutions (361.7). Entropy index indicates that the 4 classes model could explain about 51\% percent of the data, while \textit{p} values of Lo-Mendell-Rubun LRT suggest that the more classes we fit, the better model we will have until up to 6 classes (\textit{p} = 0.06 and is not shown in the table). From the parsimony point of view, we extended the model with random effects building on 2 classes, 3 classes, and 4 classes solutions.

The results of the random effect included models are presented in \textbf{Table \ref{tab:mixmodels}} under the Random effects model section. It is obvious that the BIC improves with the addition of the random effects which account for the nested structure of the data. Entropy indicates that 2 classes in the day level and 4 classes in the individual level may be the best solution mathematically. However, after these solutions were checked in more details, the potentially most substantively interpretable model was found to be the 3$\times$3 random effect model, which is the model with 3 latent classes in the day level, and 3 latent classes in the individual level. We must emphasize that different researchers may have made decision slightly different from ours, we have provided other solutions in \textbf{Appendix \ref{AppendixC}} for reference.

In the 3$\times$3 random effect model solution we have chosen, there were 39.5\%, 20.4\%, and 40.1\% observations classified into 3 latent groups at the day level. The overall counts and percentages for each response within each time slot and the distributions of the solution are presented in \textbf{Table \ref{tab:daylevel}}. The trajectories illustrating the change of the probabilities of each response to carbohydrate eating during the hours of the day are shown separately by three types of days in \textbf{Figure \ref{fig:level1}}.

@@

Class 1 days \textbf{(Figure \ref{fig:level1}-A)} were given the name of "high percentage carbohydrate day" since in these days of the survey, the probabilities of carbohydrate contributed higher or equal to 50\% of the energy consumed were always higher than that in the other two types of days. Specifically, high percentage carbohydrate days were characterised with probabilities of over 0.6 in time slots between 6 am to 9 am, 9 am to 12 noon, and also 2 pm to 5 pm, during which the time slots may be interpreted as breakfast, morning snack, and afternoon snack time periods for many participants. Moreover, even during late night time period, such as 8 pm to 10 pm, and 10 pm to 6 am time slots, the probabilities of having higher carbohydrate contained food were still as high as 0.412, and 0.246, respectively.

@@

Class 2 days \textbf{(Figure \ref{fig:level1}-B)} were named as "low percentage carbohydrate day" because first of all, in these days the probability of participants skipping breakfast was 0.45. And after 9 am, within these days, the probability of having low carbohydrate contained food (carbohydrate contributed $<$ 50\% of total energy intake), was always higher than having high carbohydrate contained food (carbohydrate contributed $\geqslant$ 50\% of total energy intake). In class 2 days, participants also turned to have morning snacks (with probability of only 0.079 of \textbf{not} eating any food and similar probabilities of having either high or low carbohydrate contained food). This phenomenon may also be interpreted as having a long and late breakfast (brunch) in these mornings. The probability of \textbf{not} eating any food was the lowest for low carbohydrate days during the midnight time slot (10 pm to 6 am), with a probability of 0.590 compared with 0.680 and 0.751 in class 1 and class 3 days, respectively.

@@

As expected, consumption of total energy (7539.98 kJ), total carbohydrate (222.79 g), total sugar (98.63 g), starch (124.07 g), and non-milk extrinsic sugar (59.45 g) were the highest among high percentage carbohydrate days (all \textit{p} < 0.001). On the other hand, the consumption of protein (73.29 g), total fat (67.24 g), and alcohol (13.80 g) were the highest in the regular meals days. Moreover, in the high percentage carbohydrate days, participants turned to consume the highest amount of fruit (107.40 g). There was no evidence of any difference in the consumption of yellow, red, or green vegetables across the three types of days (\textit{p} = 0.681).\vspace{-0.6cm}

@@

\subsection{Individual-level LCA solution}\vspace{-0.3cm}

In the random effect models, we utilized the non-parametric approach, in which we added a level 2 (individual level) latent classes based on the random means from the level 1 (day level) latent class solution. The results of the individual level LCA solution for 2 and 3 classes are presented in \textbf{Figure \ref{fig:CB2level2}}, and \textbf{\ref{fig:level2}}.

@@

With two individual-level latent classes \textbf{(Figure \ref{fig:CB2level2})}, one individual class is comprised of individuals with a relatively slightly higher proportion of having "low carbohydrate day" (22.1\%) compared to the other (17.4\%). This class represents nearly 65\% of the individuals. However, we believe these individual classes are not very distinguishable from each other.

With three individual-level latent classes \textbf{(Figure \ref{fig:level2})}, a low-carbohydrate eaters class, a moderate-carbohydrate eaters class, and a high-carbohydrate eaters class emerge. 43.1\% participants were identified as high-carbohydrate eaters, in these individuals, about 50\% of the days (2 out of 4 days) of their dietary diary could be classified as having high percentage carbohydrate days. Nearly 1 out of 4 days of their dietary diary were either "regular meals day" or "low carbohydrate day". 28.1\% participants fell into the low carbohydrate eaters class on the left-hand side of \textbf{Figure \ref{fig:level2}}, their recordings of food intake showed that in more than 60\% of their days, they were having "regular meals" which was characterised as with the highest amount of fat and alcohol consumptions as already described in \textbf{Table \ref{tab:day-level-features}}. Moderate carbohydrate eaters have comparable proportions (42.0\% vs. 40.0\%) of having high carbohydrate days and regular meals day, 18.0\% of their dietary diary was found to be low carbohydrate days.

@@

After recognising that there were three potential latent groups of carbohydrate eaters in the UK adults, whose food consumption patterns were also probably switching from one to another during the survey, their average carbohydrate contribution to total energy intake (as well as the subtypes of carbohydrate actually consumed) within the 7 pre-defined time slots of the day were still of interest. Survey-design-weighted mean energy intake within each time slot of the day and their composition of contribution is illustrated in \textbf{Figure \ref{fig:energysourcesCB}}, weighted mean nutrients intakes are listed in \textbf{Table \ref{tab:tab-nutri-indi}}.

@@

Among the three types of carbohydrate eaters, the mean of total energy intake over the 4 days of dietary survey was the highest (7985.8 kJ, 95\%CI: 7283.3, 8146.3) in the low carbohydrate eaters group, and the lowest (7341.8 kJ, 95\%CI: 7172.5, 7511.2) in the moderate eaters group \textbf{(Table \ref{tab:tab-nutri-indi})}. Sources of energy for each type of carbohydrate eaters by the 7-time slots were also different. Low carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-A)} never had carbohydrate contributed more than 50\% of their total energy throughout the day. Energy from fat was the highest for low carbohydrate eaters most of the time during the day (except for the time slot between 10 pm to 6 am next morning). Most impressively, energy contributed by alcohol was always the highest in low carbohydrate eaters, percentages for energy from alcohol for the 7-times slots were 1.4\% (6-9 am), 1.8\% (9-12 noon), 3.1\% (12-2 pm), 7.2\% (2-5 pm), 8.4\% (5-8 pm), 20.0\% (8-10 pm), and 31.2\% (10-6 am), respectively. Contributions from different energy sources were quite similar for moderate and high carbohydrate eaters, but their absolute amount of energy consumption at each time slot was largely different. Moderate carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-B)} were characterised as consuming the lowest energy (296.9 kJ) before 9 am, but having higher energy consumption (967.9 kJ) between 9 am and 12 noon time compared with low and high carbohydrate eaters. Moderate carbohydrate eaters may turn to have later breakfast, later lunch, and probably later dinner as well. They had the highest total energy consumption (599.7 kJ) at night (10 pm - 6 am) across three types of eaters. High carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-C)} consumed the highest total energy (929.0 kJ) during 6 am to 9 am in the morning and the lowest total energy between 10 pm to 6 am (205.5 kJ). Specifically, carbohydrate contributions to total energy intake were 65.3\% (6-9 am), 56.2\% (9-12 noon), 46.2\% (12-2 pm), 50.3\% (2-5 pm), 42.5\% (5-8 pm), 44.1\% (8-10 pm), and 49.9\% (10-6am). We also noticed that high carbohydrate eaters consumed their energy mainly from the following three time slots: 6-9 am, 12-2 pm, and 5-8 pm.

In total, the mean of total carbohydrate intake was 203.8 g, 218.3 g, and 233.4 g for low, moderate, and high carbohydrate eaters, respectively \textbf{(Table \ref{tab:tab-nutri-indi})}. Total energy contributed by carbohydrate was close to 50\% among the high carbohydrate eaters but was only 40.6\% among the low carbohydrate eaters. In terms of the subtypes (components) of the carbohydrate consumed at each time slot, high carbohydrate eaters consumed as much as more than 2 times (compared to low carbohydrate eaters) and nearly 4 times (against moderate carbohydrate eaters) the amount of sugar (37.9g 95\% CI: 36.8, 39.2) and non-milk extrinsic sugar (i.e. free sugar) (11.1g 95\%CI: 10.7, 11.6) between 6-9 am. Moderate carbohydrate eaters had their carbohydrate intake more spread out. They consumed more sugar and starch during 9-12 noon, 2-5 pm, 8-10 pm, and 10-6 am. Low carbohydrate eaters turned to have similar temporal pattern of consuming carbohydrates with the high carbohydrate eaters, but the absolute amount of fibre, sugar, free sugar, and starch were usually lower than that in the high carbohydrate eaters except for time slots of 2-5 pm, and 10-6 am. Strong evidence (\textit{p} < 0.001) suggested that the mean of to-

@@

tal fibre consumption for low, moderate, and high carbohydrate eaters were different: 13.7g (13.4, 14.0), 12.5g (12.1, 12.9), and 14.7g (14.4, 14.9) with all 95\% CI being exclusive to each other. It is also noteworthy that low carbohydrate eaters consumed the highest average amount of protein (79.9 g, 17.2\% of total energy), fat (74.7g, 35.4\% of total energy), and alcohol (20.8 g, 6.8\% of total energy) as we have described for \textbf{Figure \ref{fig:energysourcesCB}}.

The social-demographic characteristics of the UK adults according to their individual level latent class membership are shown in \textbf{Table \ref{tab:Level2tab1}}. Moderate carbohydrate eaters were relatively younger (\textit{p} < 0.001). Gender distribution across the three types of carbohydrate eaters was fairly even (\textit{p} = 0.119). The distribution of the carbohydrate eater types turned out to be changing with the year of survey. Low carbohydrate eaters represented 32.5\% of the population in the first year of survey but later dropped to lower than 30\% (lowest in the third year, 22.6\%) until the most recent year. The proportion of high carbohydrate eaters increased from 41.2\% to the highest (50.6\%) in the second year, but then started to decline to 38.4\% in the 8th year of survey (\textit{p} = 0.015). There was no evidence of any difference in employment status across three types of carbohydrate eaters. However, strong evidence suggested that high carbohydrate eaters had the highest proportion (61.3\%) of living with partner (\textit{p} < 0.001); moderate carbohydrate eaters had the lowest average income (27180.8 \textsterling/year), highest proportion of non-white population (20.5\%), and lower education level (23.3\% with degree of higher education) compared with either low or high carbohydrate eaters.

@@

\caption{\label{tab:Level2tab1}Weighted means, percentages, and
95\% CIs of the social-demographic characteristics by carbohydrate eating latent class memberships in the UK
adults. \\ (NDNS RP 2008/09-15/16, sample size = 6155)} \centering\vspace{-0.3cm}
\fontsize{9}{11}\selectfont

\begin{tabular}[t]{lcccc}
	\hiderowcolors
	\toprule
	\textbf{Variables} & \textbf{\Centerstack{Low carbo-\\hydrate eaters\\(n = 1730)}} & \textbf{\Centerstack{Moderate carbo-\\hydrate eaters\\(n = 1772)}} & \textbf{\Centerstack{High carbo-\\hydrate eaters\\(n = 2653)}} & \textbf{\textit{P} value} \textsuperscript{*}\\
	\midrule
	\showrowcolors
	Total (\%) & 28.3 (26.9, 29.9) & 28.7 (27.1, 30.3) & 43.0 (41.3, 44.7) & \\
%	Country (\%) &  &  &  & 0.007\\
%	\hspace{1em}England & 84.5 (81.7, 86.9) & 82.0 (79.3, 84.5) & 84.7 (82.3, 86.8) & \\
%	\hspace{1em}Northern Ireland & 2.1 (1.6, 2.8) & 4.2 (3.2, 5.6) & 2.2 (1.7, 3.0) & \\
%	\hspace{1em}Scotland & 9.1 (7.0, 11.8) & 8.6 (6.7, 11.1) & 8.0 (6.3, 10.2) & \\
%	\hspace{1em}Wales & 4.3 (3.3, 5.6) & 5.1 (4.0, 6.4) & 5.1 (4.0, 6.4) & \\
	Age (years) & 51.0 (49.9, 52.1) & 40.3 (39.1, 41.6) & 51.7 (50.7, 52.7) & < 0.001\\
	Sex (\%) &  &  &  & 0.119\\
	\hspace{1em}Men & 50.0 (46.9, 53.1) & 50.2 (47.0, 53.5) & 46.6 (44.0, 49.1) & \\
	\hspace{1em}Women & 50.0 (46.9, 53.1) & 49.8 (46.5, 53.0) & 53.4 (50.9, 56.0) & \\
	Survey years (\% in rows) &  &  &  & 0.015\\
	\hspace{1em}1 & 32.5 (28.4, 36.9) & 26.3 (21.9, 31.2) & 41.2 (36.6, 46.0) & \\
	\hspace{1em}2 & 26.8 (22.6, 31.3) & 22.6 (18.6, 27.3) & 50.6 (45.8, 55.4) & \\
	\hspace{1em}3 & 22.6 (18.8, 26.9) & 33.7 (28.6, 39.2) & 43.6 (38.7, 48.7) & \\
	\hspace{1em}4 & 27.9 (24.1, 32.2) & 27.6 (23.8, 31.8) & 44.4 (40.2, 48.7) & \\
	\hspace{1em}5 & 27.9 (24.2, 32.0) & 28.7 (24.4, 33.5) & 43.3 (38.2, 48.6) & \\
	\hspace{1em}6 & 28.0 (24.0, 32.4) & 31.5 (26.9, 36.6) & 40.5 (35.8, 45.3) & \\
	\hspace{1em}7 & 29.1 (25.2, 33.4) & 29.0 (24.5, 34.0) & 41.8 (37.1, 46.7) & \\
	\hspace{1em}8 & 31.1 (27.3, 35.3) & 30.5 (25.9, 35.5) & 38.4 (34.1, 42.8) & \\
	Paid employment\textsuperscript{\dag} (\%) &  &  &  & 0.907\\
	\hspace{1em}Yes & 40.3 (37.0, 43.6) & 40.8 (37.1, 44.5) & 39.8 (37.1, 42.6) & \\
	\hspace{1em}No & 59.7 (56.4, 63.0) & 59.2 (55.5, 62.9) & 60.2 (57.4, 62.9) & \\
	Live with partner\textsuperscript{\ddag} (\%) &  &  &  & < 0.001\\
	\hspace{1em}Yes & 56.9 (53.6, 60.1) & 38.4 (35.2, 41.8) & 61.3 (58.7, 63.7) & \\
	\hspace{1em}No & 43.1 (39.9, 46.4) & 61.6 (58.2, 64.8)) & 38.7 (36.3, 41,3) & \\
	Household income, \textsterling/year & \Centerstack{36558.5\\(34800.2, 38316.8)} & \Centerstack{27180.8\\(25597.9, 28763.7)} & \Centerstack{32171.6\\(31024.9, 33318.2)} & < 0.001\\
	Ethnicity (\%) &  &  &  & \\
	\hspace{1em}White & 94.2 (92.4, 95.6) & 79.5 (76.4, 82.3) & 91.9 (90.1, 93.4) & < 0.001\\
	\hspace{1em}Non-White & 5.8 (4.4, 7.6) & 20.5 (17.7, 23.6) & 8.1 (6.6, 9.9) & \\
	Education (\%) &  &  &  & \\
	\hspace{1em}Degree or higher & 29.0 (26.1, 32.1) & 23.3 (20.5, 26.3) & 26.2 (24.1, 28.5) & 0.019\\
	\hspace{1em}Lower than degree & 71.0 (67.9, 73.9) & 76.7 (73.7, 79.5) & 73.8 (71.5, 75.9) & \\
	\bottomrule
	\multicolumn{5}{l}{{\scriptsize \textit{Note: }}}\\
	\multicolumn{5}{l}{{\scriptsize \textbf{Abbreviations}: CI, confidence intervals; NDNS RP, national diet and nutrition survey rolling programme.}}\\
	\multicolumn{5}{l}{{\scriptsize Variables were weighted by individual weights.}}\\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{*} For continuous variables, the \textit{F} test was used to determine differences between latent classes with Bonferroni}}\\ 
	\multicolumn{5}{l}{{\scriptsize correction to account for multiple testing across $>$ 2 classes. For categorical variables, differences between}}\\
	\multicolumn{5}{l}{{\scriptsize  latent classes were assessed using the adjusted Pearson $\chi^2$ test for survey data.}} \\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{\dag} Paid employment was defined as being in paid employment during the last 4 weeks prior to the survey.} }\\
	\multicolumn{5}{l}{{\scriptsize \textsuperscript{\ddag} Live with partner was defined as either living with a married husband/wife or a legally recognised civil}} \\
	\multicolumn{5}{l}{{\scriptsize partnership.}}\\
\end{tabular}

@@

\subsection{Association between individual-level latent classes and hypertension, and obesity.}\vspace{-0.3cm}
